---
title: "Mental Health and Suicide Trends Analysis (2002–2022)"
output: html_document
---

## 1. Descriptive Analysis 

- How have **mental health indicators** changed over time (**2002, 2012, 2022**) across different demographics (age, province, sex) at both **individual and provincial levels**?
- What are the **suicide rate trends in Canada** over the 20-year period?  
  - How do suicide rates vary across different **demographic groups (age, sex, ethnicity, income level)?**  
  - How do suicide rates vary at the **province level**? Are certain provinces disproportionately affected?
- Are there **regional disparities** in mental health conditions and suicide rates?  
  - Which **provinces have the highest vs. lowest rates** of mental health disorders and suicide?  
- How have **economic indicators (CPI, inflation, unemployment)** varied across provinces over time?  
  - Does economic distress correlate with worsening mental health trends?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
```

```{r load-data}
df <- read.csv("C:/Users/felix/Downloads/cleaned_mental_health_data.csv", header=TRUE)
View(df)
```

```{r exploratory-analysis}

summary(df)
glimpse(df)
```

```{r visualization-example}

```

