---
title: "Hypothesis Testing"
author: "Felix Vo"
date: "2025-03-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#Import libraries:
library(tidyverse)
library(knitr)
```

# Hypothesis Testing:

-   **Statistical relationship between mental health disorder and number of commit suicide**

```{r mental health data, echo=FALSE}
mental_health_data <- read.csv("C:/Users/felix/Desktop/CODING/felix's works/Mental-Health-On-Suicide-Rates-Trend-Analysis-Prediction/datasets/processed/merged_suicide_mental_health_data.csv")
```

```{r}

df_canada <- mental_health_data %>%
  filter(Geography == "Canada", Gender == "Total, gender of person")
suicide_indicator <- "Commit Suicide"

mental_health_indicators <- c("Major depressive episode, life", 
                              "Eating disorder, current diagnosed condition",
                              "Suicidal thoughts, life", 
                              "Suicidal thoughts, 12 months",
                              "Social phobia, life", 
                              "Generalized anxiety disorder, life",
                              "Bipolar disorder, life")

df_filtered <- df_canada %>%
  filter(Indicators %in% c(suicide_indicator, mental_health_indicators)) %>%
  select(Year, Indicators, Population) %>%
  group_by(Year, Indicators) %>%
  summarise(Population = sum(Population, na.rm = TRUE), .groups = "drop")

df_pivot <- df_filtered %>%
  pivot_wider(names_from = Indicators, values_from = Population)

print(df_pivot)

```

```{r}
# Handle missing values (replace NA with median)
df_pivot_filled <- df_pivot %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

# Correlation analysis
cor_results <- data.frame(
  Indicator = character(),
  Correlation = numeric(),
  t_value = numeric(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)

for (indicator in mental_health_indicators) {
  if (indicator %in% colnames(df_pivot_filled)) {
    cor_test <- cor.test(df_pivot_filled[[indicator]], df_pivot_filled[["Commit Suicide"]])
    cor_results <- rbind(cor_results, data.frame(
      Indicator = indicator,
      Correlation = round(cor_test$estimate, 3),
      t_value = round(cor_test$statistic, 3),
      p_value = round(cor_test$p.value, 4)
    ))
  } else {
    cor_results <- rbind(cor_results, data.frame(
      Indicator = indicator,
      Correlation = NA,
      t_value = NA,
      p_value = NA
    ))
  }
}

# Display results in a table
kable(cor_results, caption = "Pearson Correlation Between Mental Health Indicators and Suicide Rates")
```

## Suicide Thoughts with other Indicators Pearson correlation test

```{r}
suicide_indicators <- c("Suicidal thoughts, life", "Suicidal thoughts, 12 months")

mental_health_indicators <- c("Major depressive episode, life", 
                              "Eating disorder, current diagnosed condition",
                              "Social phobia, life", 
                              "Generalized anxiety disorder, life",
                              "Bipolar disorder, life")

df_filtered <- df_canada %>%
  filter(Indicators %in% c(suicide_indicators, mental_health_indicators)) %>%
  select(Year, Indicators, Population) %>%
  group_by(Year, Indicators) %>%
  summarise(Population = sum(Population, na.rm = TRUE), .groups = "drop")  

df_pivot <- df_filtered %>%
  pivot_wider(names_from = Indicators, values_from = Population)

df_pivot_filled <- df_pivot %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))

cor_results <- data.frame(
  Indicator = character(),
  Suicide_Indicator = character(),
  Correlation = numeric(),
  t_value = numeric(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)

for (suicide in suicide_indicators) {
  for (indicator in mental_health_indicators) {
    if (indicator %in% colnames(df_pivot_filled) && suicide %in% colnames(df_pivot_filled)) {
      cor_test <- cor.test(df_pivot_filled[[indicator]], df_pivot_filled[[suicide]])
      cor_results <- rbind(cor_results, data.frame(
        Indicator = indicator,
        Suicide_Indicator = suicide,
        Correlation = round(cor_test$estimate, 3),
        t_value = round(cor_test$statistic, 3),
        p_value = round(cor_test$p.value, 4)
      ))
      
      print(paste("Correlation between", indicator, "and", suicide, ":"))
      print(cor_test)
      cat("\n")
    } else {
      cor_results <- rbind(cor_results, data.frame(
        Indicator = indicator,
        Suicide_Indicator = suicide,
        Correlation = NA,
        t_value = NA,
        p_value = NA
      ))
    }
  }
}

kable(cor_results, caption = "Pearson Correlation Between Mental Health Indicators and Suicidal Thoughts")
```

## Summary Table

| **Mental Health Indicator** | **Suicide Indicator** | **Correlation (r)** | **t-value** | **p-value** | **Significance** |
|------------|------------|------------|------------|------------|------------|
| **Major depressive episode, life** | Suicidal thoughts, life | **0.9997** | 38.093 | **0.0167** | **Significant** |
| **Eating disorder, current diagnosed condition** | Suicidal thoughts, life | 0.9880 | 6.3971 | 0.0987 | Not significant |
| **Social phobia, life** | Suicidal thoughts, life | NA | NA | NA | Not available |
| **Generalized anxiety disorder, life** | Suicidal thoughts, life | 0.8236 | 1.452 | 0.384 | Not significant |
| **Bipolar disorder, life** | Suicidal thoughts, life | 0.8236 | 1.452 | 0.384 | Not significant |
| **Major depressive episode, life** | Suicidal thoughts, 12 months | **0.9999** | 83.325 | **0.0076** | **Significant** |
| **Eating disorder, current diagnosed condition** | Suicidal thoughts, 12 months | 0.9857 | 5.8497 | 0.1078 | Not significant |
| **Social phobia, life** | Suicidal thoughts, 12 months | NA | NA | NA | Not available |
| **Generalized anxiety disorder, life** | Suicidal thoughts, 12 months | 0.8154 | 1.4086 | 0.393 | Not significant |
| **Bipolar disorder, life** | Suicidal thoughts, 12 months | 0.8154 | 1.4086 | 0.393 | Not significant |

------------------------------------------------------------------------

```{r}
economic_data  <-read.csv("C:/Users/felix/Desktop/CODING/felix's works/Mental-Health-On-Suicide-Rates-Trend-Analysis-Prediction/datasets/processed/CPI_inflation.csv")
```
```{r}
economic_canada <- economic_data %>%
  filter(Geo == "Canada") %>%
  select(Year, CPI, Inflation_rate)

suicide_indicators <- c("Suicidal thoughts, life", "Suicidal thoughts, 12 months")

suicide_canada <- df_canada %>%
  filter(Geography == "Canada", Gender == "Total, gender of person", Indicators %in% suicide_indicators) %>%
  select(Year, Indicators, Population) %>%
  group_by(Year, Indicators) %>%
  summarise(Population = sum(Population, na.rm = TRUE), .groups = "drop")

# Convert suicide data to wide format
suicide_pivot <- suicide_canada %>%
  pivot_wider(names_from = Indicators, values_from = Population)

# Merge suicide and economic data
merged_data <- merge(suicide_pivot, economic_canada, by = "Year", all.x = TRUE)

# Display merged data
kable(merged_data, caption = "Merged Economic and Suicide Data in Canada")
```


```{r}
# Correlation Analysis
cor_results <- data.frame(
  Suicide_Indicator = character(),
  Economic_Indicator = character(),
  Correlation = numeric(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)

economic_indicators <- c("CPI", "Inflation_rate")

for (suicide in suicide_indicators) {
  for (economic in economic_indicators) {
    if (suicide %in% colnames(merged_data) && economic %in% colnames(merged_data)) {
      cor_test <- cor.test(merged_data[[suicide]], merged_data[[economic]])
      cor_results <- rbind(cor_results, data.frame(
        Suicide_Indicator = suicide,
        Economic_Indicator = economic,
        Correlation = round(cor_test$estimate, 3),
        p_value = round(cor_test$p.value, 4)
      ))
      
      print(paste("Correlation between", suicide, "and", economic, ":"))
      print(cor_test)
      cat("\n")
    }
  }
}

# Display correlation results
kable(cor_results, caption = "Correlation Between Economic Indicators and Suicidal Thoughts")
```

#### Inflation Rate has a strong relationship with suicidal thoughts:

Inflation rate is highly correlated with:
Suicidal thoughts, life (r = 0.9759, p = 0.1400)
Suicidal thoughts, 12 months (r = 0.9727, p = 0.1491)
The p-values are slightly above 0.05, meaning the results are not statistically significant.

#### CPI is not a strong predictor:

CPI (Consumer Price Index) has a moderate correlation with suicidal thoughts, but the p-values are too high to conclude a meaningful relationship.
